### 2018-05-29 mmm-minorcycle.tex
Sub image MAM memo as submitted

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Distribution of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-CASS-001\xspace}
% Context
\newcommand{\context}{(ASKAPsoft)}
% Revision
\newcommand{\revision}{0.1\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Released\xspace}

@@

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
    0.1 & 03/04/2018 & T.J. Cornwell & Initial version \\
    \hline
  \end{tabularx}
}

@@

% Add here the abbreviations used in your document
\sdplistofabbreviations
\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
    \item[ASKAP] Australian Square Kilometer Array Pathfinder\vspace{-0.2cm}
    \item[ARL] Algorithm Reference Library\vspace{-0.2cm}
    \item[FFT] Fast Fourier Transform\vspace{-0.2cm}
    \item[FLOP] Floating Point Operations \vspace{-0.2cm}
    \item[FLOPS] Floating Point Operations per Second \vspace{-0.2cm}
    \item[FoV] Field of View\vspace{-0.2cm}
    \item[MS-MFS] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[MAM] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[PSF] Point Spread Function \vspace{-0.2cm}
    \item[SDP] Scientific Data Processor\vspace{-0.2cm}
    \item[SKA] Square Kilometre Array\vspace{-0.2cm}
%    \item[] \vspace{-0.2cm}
\end{basedescript}

@@

We investigate the various ways that the distribution can be performed.

@@

 \begin{center}
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
   {\bf [RD01]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
   {\bf [RD02]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
   {\bf [RD03]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
   {\bf [RD04]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
   {\bf [RD05]} & Junklewitz, H., Bell, M. R., and Ensslin, T. (2015). A new approach to multifrequency synthesis in radio interferometry. Astronomy \& Astrophysics, 581, A59. http://doi.org/10.1051/0004-6361/201423465 \\
	{\bf [RD06]} & Rau, U., \& Cornwell, T. J. (2011, June 14). A multi-scale multi-frequency deconvolution algorithm for synthesis imaging in radio interferometry. arXiv.org. EDP Sciences. http://doi.org/10.1051/0004-6361/201117104   \\
	{\bf [RD07]} & Bhatnagar, S., Rau, U., \& Golap, K. (2013). Wide-field wide-band Interferometric Imaging: The WB A-Projection and Hybrid Algorithms. The Astrophysical Journal, 770(2), 91. http://doi.org/10.1088/0004-637X/770/2/91 \\
	{\bf [RD08]} & Cornwell, T.J. (2016), Implementation of the Rau-Cornwell MSMFS algorithm, TCC-SDP-151123-2. \\
	\hline
 \end{tabularx}
\end{center}

@@

The purpose of this document is to describe options for distributing the minor cycle of the Multi-Scale MultiFrequency Synthesis algorithm, as implemented in ASKAPsoft,.

@@

The scope includes only the published Multi-Scale, Multi-Frequency algorithms, existing implementations, and inclusion in the ASKAPsoft processing pipelines. Only the minor cycle of that algorithm is addressed. In addition, the many possible threading possibilities are not explored.

@@

The \MAM\ algorithm implementation has been discussed extensively in RD08. We use the same terminology and symbols.

The (serial) minor cycle from the Urvashi-Cornwell algorithm [RD03, RDO6] is shown below in both pseudo-code (Algorithm \ref{alg:minor_cycle_UC2011}) and python (Figure \ref{fig:minor_cycle_python}).

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./processing_histogram.png}
  \caption{Estimation of the cost of processing steps from the SKA Performance Model as adapted to ASKAP. Left column is for large PSF patch size, right is for patch size of 512, which is the recommended value for ASKAP processing.}
  \label{fig:processing_histogram}
\end{figure}

@@

The fraction of processing, $p$, for \MAM\ is about 4 \%. Thus the maximum possible speedup of the ICAL pipeline is about 25.
\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./AmdahlsLaw.png}
  \caption{Amdahl's law for various fractions of parallel portion.}
  \label{fig:amdahl}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./tapers.png}
  \caption{Tapers for feathering facet images. The vertical pink lines show the where the overlaps start. This is for 256 pixels and overlap of 64. For this case, the efficiency is roughly $(64/256)^2 = 1/16$}
  \label{fig:tapers}
\end{figure}

\subsection{Load balancing}

The primary cause of load in-balance is that different sub-images will have different numbers of sources and hence different elapsed time for deconvolution. In the graph-based approach used in the \ARL\ this is somewhat mitigated by the intrinsic flexibility of the graph-based work allocation. In MPI, it is possible that some explicit load-balancing may be required. We therefore provide some analysis of the need for load balancing.

We can improve on this rough argument by some straightforward modeling. To a reasonable approximation, the processing time for a given sub-image will be proportional to the fraction of a flux in that sub-image. For a given model of the sky, we can calculate the fraction of time taken for each sub-image\footnote{See section \ref{sec:simulations} for more details of the simulations}. We illustrate that point in figures \ref{fig:fractions_without_pb}, \ref{fig:fractions_with_pb}, and \ref{fig:fractions_clipped_image}. The first plot is a realistic astronomical field constructed from the S3-SEX catalog, selected for all sources brighter than 10uJy/beam. The curve shows the fraction of flux per sub-image against rank of that flux. The slope comes from the logN/logS statistics of the S3-SEX catalog. The peak value shows roughly the processing for the worst sub-image. This provides an upper bound on the scaling possible. The variable nproc is an estimate for the number of processors derived by dividing the total amount of processing by the peak value needed.

The next plot (figure \ref{fig:fractions_with_pb}) shows the effect of the ASKAP (single) primary beam. Most sub-images do not require cleaning and all the processing goes to the brightest fields. As processing proceeds thorough a number of major cycles, the brightest sources are been removed and eventually there remains a sea of sources spread other the primary and into the primary beam side-lobes. At this point, we can model the flux as simply by clipping the image with the primary beam at some maximum brightness such as 1mJy. The resulting scaling shown in figure \ref{fig:fractions_clipped_image} is then better balanced.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./fluxes_without_pb_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) without a primary beam applied. Scales are [0, 3, 10], and the overlap is 16 pixels. The slope for all curves is mainly due to the statistics of image which is derived from radio source LogN/LogS relationship.}
  \label{fig:fluxes_without_pb}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./fluxes_with_pb_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) with a primary beam applied. Scales are [0, 3, 10], and the overlap is 16 pixels. Compared to figure \ref{fig:fluxes_without_pb}}, the primary beam application shifts the high rank sub-images lower in flux.
  \label{fig:fluxes_with_pb}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./fluxes_clipped_image_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank for various numbers of sub-images, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10], and the overlap is 16 pixels. There are now no bright sources left.}
  \label{fig:fractions_clipped_image}
\end{figure}

To get the overall scaling curve we must sum the work required for a sequence of major cycles going successively deeper. The thresholds start at 0.1 Jy and diminish by $\sqrt(10)$ as we proceed through 10 major cycles. We sum the results for each threshold weighting according to the number of sources at that level. We perform this analysis for both 2km and 6km configurations. The results are shown in figure \ref{fig:aggregate_scaling_2km} and \ref{fig:aggregate_scaling_6km}. The scaling is driven by the larger number of weak sources. The flatness of the curves as a function of rank means that strong scaling is possible for 200 - 300 processors. Adding more processors beyond this limit brings diminishing returns, and the overall efficiency drops.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./aggregate_scaling_ASKAP_2km.png}
  \caption{Aggregate scaling curve for 2km ASKAP configuration. Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10], and the overlap is 16 pixels.}
  \label{fig:aggregate_scaling_2km}
\end{figure}

We have repeated this calculation for the full 36 antenna 6km ASKAP (see figure \ref{fig:aggregate_scaling_6km}).

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./aggregate_scaling_ASKAP_6km.png}
  \caption{Aggregate predicted scaling curve for 6km configuration. Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10, 30], and the overlap is 48 pixels.}
  \label{fig:aggregate_scaling_6km}
\end{figure}

@@

We made simulations as follows:
\begin{itemize}
	\item ASKAP antennas in 2km configuration, 
	\item 11 frequencies spread from 0.85GHz to 1.15GHz, enabling fitting of  Taylor series up to 5 terms (quintic). The bandwidth is sufficiently broad that over the frequency range far out sources can appear and disappear at different frequencies. Hence 5 terms are used for many investigations.
	\item The model was generated by querying the S3-SEX database\footnote{http://s-cubed.physics.ox.ac.uk/s3\_sex}, with a flux cutoff at 10uJy/beam. The spectral index was derived from the 1400MHz and 610MHz fluxes.
	\item Integrations spaced at 12 minutes, no noise added,
	\item Field center at ra=+30deg, dec=-60deg,
	\item Primary beam of 12m antenna with 2m blockage, no feed-legs.
	\item For some simulations the w term was set to zero to concentrate on the cleaning timing.
\end{itemize}

Figure \ref{fig:ASKAP_sim_restored} shows a typical restored image obtained by running 5 major cycles MSMFS with 3 Taylor terms. The residual image (\ref{fig:ASKAP_sim_residual}) shows emission out into the first lobe of the primary beam.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./ASKAP_sim_restored.png}
  \caption{Test field imaged with the MSMFS, major/minor cycle algorithm in the ARL. The fit is to the first three Taylor terms (i.e. quadratic). No faceting has been used for this image and the PSF support is to the entire field.}
  \label{fig:ASKAP_sim_restored}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./ASKAP_sim_residual.png}
  \caption{Residual for test field imaged with the MSMFS, major/minor cycle algorithm in the ARL. The fit is to the first three Taylor terms (i.e. quadratic). No faceting has been used for this image and the PSF support is to the entire field.}
  \label{fig:ASKAP_sim_residual}
\end{figure}

@@

@@

For a single iteration in the minor cycles, the processing breaks down into a search, scaling as the number of pixels, and a subtraction scaling as the product of the number of pixels in the image and the size of the PSF support. The value of the PSF support used thus directly affects the processing time. Furthermore, the processing via sub-image inevitably means that a limit PSF support will be used.

For the simulated case, the RMS side-lobes of the synthesized beam are about 0.3\% or less for pixels more than a few beams from the PSF peak. Hence we have reason to be optimistic that small values of PSF support will be acceptable given that we use a major/minor cycle. It is therefore important to determine the relationship is between final image quality and the PSF support.

We investigated the impact of limit PSF support using the simulations described above. In figure \ref{fig:error_vs_support}, we show the fractional error between the images obtained with the restricted PSF and the full PSF as a function of the support. The curve shows that the the effects of limited PSF support are much less than the criterion to switch to a major cycle (typical 10\% or less).

Clearly, small PSF support should be used: the speedup is significant and there is no significant increase in error. As a corollary, the use of sub-images in the deconvolution should be acceptable within a major/minor cycle algorithm.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./error_versus_support.png}
  \caption{Fractional discrepancy as a function of PSF support. This shows a parallel 2048 by 2048 pixel, 3 term MSMFS clean in ARL. The computer was running 4 Dask workers with 32GB each. The discrepancy is the difference between the image processed with PSF support and the image processed with PSF support set to cover the entire image.}
  \label{fig:error_vs_support}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./time_versus_support}
  \caption{Processing time (seconds) as a function of PSF support. This shows  a parallel 1024 by 1024 pixel, 3 term MSMFS clean in ARL. The computer was running 4 Dask workers with 32GB each.}
  \label{fig:time_vs_support}
\end{figure}

\subsection{Distribution of residuals}

Do the residuals for no overlap and overlap have similar distributions?

\subsection{Spatial structure induced by partitioning}

Do the residuals show excess power on the scale of the overlap?

\pagebreak
\section{Porting to MPI}

@@

Since the scaling curves are relatively flat (see e.g.figures \ref{fig:aggregate_scaling} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing. Load balancing would then only be needed for the high rank tail.

@@

>>>>

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Distribution of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-CASS-001\xspace}
% Context
\newcommand{\context}{(ASKAPsoft)}
% Revision
\newcommand{\revision}{0.9\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Draft for review\xspace}

@@

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
    0.9 & 30/04/2018 & T.J. Cornwell & Initial version \\
    \hline
  \end{tabularx}
}

@@

% Add here the abbreviations used in your document
\sdplistofabbreviations
\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
    \item[ASKAP] Australian Square Kilometer Array Pathfinder\vspace{-0.2cm}
    \item[ARL] Algorithm Reference Library\vspace{-0.2cm}
    \item[FFT] Fast Fourier Transform\vspace{-0.2cm}
    \item[FoV] Field of View\vspace{-0.2cm}
    \item[MS-MFS] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[MAM] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[PSF] Point Spread Function \vspace{-0.2cm}
    \item[SDP] Scientific Data Processor\vspace{-0.2cm}
    \item[SKA] Square Kilometre Array\vspace{-0.2cm}
%    \item[] \vspace{-0.2cm}
\end{basedescript}

@@

We analyze the various ways that the distribution of processing can be performed and what performance we can expect. We develop and demonstrate an algorithm. Out analysis shows that strong scaling will possible up to hundreds of nodes.

The algorithm has been coded in python using a graph-based distribution package, Dask. We describe the steps necessary to implement the algorithm in ASKAPsoft using MPI to achieve the distribution.

@@

 \begin{center}
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
     {\bf [RD01]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
     {\bf [RD02]} & U.Rau, T.J.Cornwell, A multi-scale multi-frequency deconvolution algorithm for synthesis imaging in radio-interferometry, , Astronomy and Astrophysics, Volume 532, August 2011.\\
	{\bf [RD03]} & Cornwell, T.J. (2016), Implementation of the Rau-Cornwell MSMFS algorithm, TCC-SDP-151123-2. \\
	\hline
 \end{tabularx}
\end{center}

@@

The purpose of this document is to describe options for distributing the minor cycle of the Multi-Scale MultiFrequency Synthesis algorithm, as implemented in ASKAPsoft.

@@

The scope includes only the published Multi-Scale, Multi-Frequency algorithm, existing implementations, and inclusion in the ASKAPsoft processing pipelines. Only distribution of the minor cycle of that algorithm is addressed. In addition, the many possible threading possibilities are not explored.

@@

The \MAM\ algorithm implementation has been discussed extensively in RD03. We use the same terminology and symbols.

The (serial) minor cycle from the Urvashi-Cornwell algorithm [RD01, RD02, RDO3] is shown below in both pseudo-code (Algorithm \ref{alg:minor_cycle_UC2011}) and python (Figure \ref{fig:minor_cycle_python}).

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./pngs/processing_histogram.png}
  \caption{Estimation of the cost of processing steps from the SKA Performance Model as adapted to ASKAP. Left column is for large PSF patch size, right is for patch size of 512, which is the recommended value for ASKAP processing.}
  \label{fig:processing_histogram}
\end{figure}

@@

The fraction of processing, $p$, for \MAM\ is about 4 \%. Thus the maximum possible speedup of the ICAL pipeline is about 25.
\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./pngs/AmdahlsLaw.png}
  \caption{Amdahl's law for various fractions of parallel portion.}
  \label{fig:amdahl}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{./pngs/tapers.png}
  \caption{Tapers for feathering facet images. The vertical pink lines show the where the overlaps start. This is for 256 pixels and overlap of 64. For this case, the efficiency is roughly $(64/256)^2 = 1/16$}
  \label{fig:tapers}
\end{figure}

@@

We made simulations as follows:
\begin{itemize}
	\item ASKAP antennas in 2km configuration, 
	\item 11 frequencies spread from 0.85GHz to 1.15GHz, enabling fitting of  Taylor series up to 5 terms (quintic). The bandwidth is sufficiently broad that over the frequency range far out sources can appear and disappear and then appear again at different frequencies. Hence 5 terms are used for many investigations.
	\item The source catalogs generated by querying the S3-SEX database\footnote{http://s-cubed.physics.ox.ac.uk/s3\_sex}, with a flux cutoff at 10uJy/beam and 100uJy/beam. The spectral index was derived from the 1400MHz and 610MHz fluxes.\footnote{Caveat emptor: S3-SEX contains a 730 Jy source in the catalog. We have removed this prior to these simulations on the basis that it should be dealt with by peeling instead of cleaning.}
	\item Integrations spaced at 12 minutes, no noise added,
	\item Field center at ra=+30deg, dec=-60deg,
	\item Primary beam of 12m antenna with 2m blockage, no feed-legs.
	\item For some simulations the w term was set to zero to concentrate on the cleaning timing.
	\item The subimages are calculated so that the size is constant. Thus if the overlap is non-zero, the actual number of sub-images on an axis is two less than the nominal {\em e.g.} If the nominal is 8 then the actual is 6 and then there will be 36 sub-images spread over the two dimensions. We chose this approach to avoid having to  analyze the behaviour for different size subimages.
\end{itemize}

Figure \ref{fig:ASKAP_sim_restored} shows a typical restored image obtained by running 5 major cycles \MAM\ with 5 Taylor terms. The residual image (\ref{fig:ASKAP_sim_residual}) shows emission out into the first lobe of the primary beam.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/ASKAP_sim_restored.png}
  \caption{Test field imaged with the MSMFS, major/minor cycle algorithm in the ARL. The fit is to the first five Taylor terms (i.e. quintic). 16 facets on each axis were used, with overlap 16 pixels.}
  \label{fig:ASKAP_sim_restored}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/ASKAP_sim_residual.png}
  \caption{Residual for test field imaged with the MSMFS, major/minor cycle algorithm in the ARL. The fit is to the first five Taylor terms (i.e. quintic polynomial). 16 facets on each axis were used, with overlap 16 pixels.}
  \label{fig:ASKAP_sim_residual}
\end{figure}

@@

The first possible stumbling block is that the cleaning requires a PSF that crosses facet boundaries. If this was actually correct then the sub-image \MAM\ would not be feasible.

In the current single threaded \MAM\ the minor cycles, the processing consists of a search step, scaling as the number of pixels, and a subtraction step,  scaling as the product of the number of pixels in the image and the size of the PSF support. The value of the PSF support used thus directly affects the processing time. Furthermore, the processing via sub-image inevitably means that a limit PSF support will be used.

For the simulated case, the RMS side-lobes of the (uniform weighted) synthesized beam are about 0.3\% or less for pixels more than a few beams from the PSF peak. Hence we have reason to be optimistic that small values of PSF support will be acceptable given that we use a major/minor cycle. It is therefore important to determine the relationship is between final image quality and the PSF support.

We investigated the impact of limit PSF support using the simulations described above. In figure \ref{fig:error_vs_support}, we show the fractional error between the images obtained with the restricted PSF and the full PSF as a function of the support. The curve shows that the the effects of limited PSF support are much less than the criterion to switch to a major cycle (typically 10\% or less).

Clearly, in the context of a major/minor cycle algorithm, small PSF support can be used: the speedup is significant and there is no significant increase in error compared to the cutoff for the transition from minor to major cycle. This is confirmed by evaluating the difference between images processed using \MAM\ and Sub-image \MAM\ after ten major cycles (see figure \ref{fig:error_vs_support_ci}.

We should emphasize that this result also holds for the current use of the single threaded \MAM\: the psf support can be set to a value much smaller than 512. One possible limit would be where the subtraction step becomes comparable to the search cost. This is obviously context dependent but we would expect values in the range 32 - 128. However, figure \ref{fig:time_vs_support} shows that the gains are relatively small.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/error_versus_support.png}
  \caption{Fractional discrepancy as a function of PSF support. This shows a parallel 2048 by 2048 pixel, 3 term \MAM\ clean in ARL. The computer was running 4 Dask workers with 32GB each. The discrepancy is the difference between the image processed with PSF support and the image processed with PSF support set to cover the entire image.}
  \label{fig:error_vs_support}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/error_versus_support_ci.png}
  \caption{Fractional discrepancy as a function of PSF support after five major cycles. This shows a parallel 2048 by 2048 pixel, 3 term \MAM\ clean in ARL. The computer was running 4 Dask workers with 32GB each. The discrepancy is the difference between the image processed with PSF support and the image processed with PSF support set to cover the entire image.}
  \label{fig:error_vs_support_ci}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/time_versus_support}
  \caption{Processing time (seconds) as a function of PSF support. This shows  a parallel 1024 by 1024 pixel, 3 term \MAM\ clean in ARL. The computer was running 4 Dask workers with 32GB each.}
  \label{fig:time_vs_support}
\end{figure}

@@

We turn now to the connection between the partitioning of the image by the used of overlapped facets and the performance of the sub-imaging algorithm. Accepting the performance as given by the existing single-threaded code running in parallel on many sub-images, the key question is whether the processing load is well-balance and provides strong scaling.

The primary cause of load imbalance is that different sub-images will have different numbers of sources and hence different elapsed time for deconvolution. In the graph-based approach used in the \ARL\ this is observed directly. In ARL it is somewhat mitigated by the intrinsic flexibility of the graph-based work allocation.

We can improve on this rough argument by some straightforward modeling. To a reasonable approximation, the processing time for a given sub-image will be proportional to the fraction of a flux in that sub-image. For a given model of the sky, we can calculate the fraction of time taken for each sub-image. We illustrate that point in figures \ref{fig:fluxes_without_pb}, \ref{fig:fluxes_with_pb}, and \ref{fig:fractions_clipped_image}. The first plot is a realistic astronomical field constructed from the S3-SEX catalog, selected for all sources brighter than 10uJy/beam. The curve shows the fraction of flux per sub-image against rank of that flux. The slope comes from the logN/logS statistics of the S3-SEX catalog. The peak value shows roughly the processing for the worst sub-image. This provides an upper bound on the scaling possible. The mean number of processors is an estimate derived by dividing the total amount of processing by the peak value needed.

The next plot (figure \ref{fig:fractions_with_pb}) shows the effect of the ASKAP (single) primary beam. Most sub-images do not require cleaning and all the processing goes to the brightest fields. As processing proceeds thorough a number of major cycles, the brightest sources are been removed and eventually there remains a sea of sources spread other the primary and into the primary beam side-lobes. For the final major cycles, we can model the flux as simply by clipping the image with the primary beam at some maximum brightness such as 1mJy. The resulting scaling shown in figure \ref{fig:fractions_clipped_image} is then better balanced.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/fluxes_No_primary_beam_ASKAP_2km_2048.png}
  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) without a primary beam applied. Scales are [0, 3, 10], and the overlap is 16 pixels. The slope for all curves is mainly due to the statistics of image which is derived from radio source LogN/LogS relationship. This shows the complete absence of strong scaling i.e. adding more processors does not bring commensurate improvements in performance.}
  \label{fig:fluxes_without_pb}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/fluxes_with_pb_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank, for realistic image (constructed from S3-SEX) with a primary beam applied. Scales are [0, 3, 10], and the overlap is 16 pixels. Compared to figure \ref{fig:fluxes_without_pb}, the primary beam application shifts the high rank sub-images lower in flux. The scaling has improved slightly.}
  \label{fig:fluxes_with_pb}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/fluxes_clipped_image_ASKAP_2km.png}
  \caption{Integrated sub-image flux versus rank for various numbers of sub-images, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 100uJy/beam. Scales are [0, 3, 10], and the overlap is 16 pixels.} 
  \label{fig:fractions_clipped_image}
\end{figure}

To get the overall scaling curve we must sum the work required for a sequence of major cycles going successively deeper. The thresholds start at $\sqrt{0.1}$ Jy and diminish by $\sqrt{10}$ as we proceed through 10 major cycles. We sum the results for each threshold. We perform this analysis for both 2km and 6km configurations. The results are shown in figure \ref{fig:aggregate_scaling_2km} and \ref{fig:aggregate_scaling_6km}. The scaling is driven by the larger number of weak sources. The flatness of the curves as a function of rank means that strong scaling is possible for 200 - 300 processors. Adding more processors beyond this limit brings diminishing returns, and the overall efficiency drops.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/aggregate_scaling_single_ASKAP_beam_ASKAP_2km_2048.png}
  \caption{Aggregate scaling curve for 2km ASKAP configuration on multiple major cycles. Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10], and the overlap is 16 pixels. The image size is 2048 by 2048 pixels. The scaling of work with numbers of sub-images (i.e. independent processors) is much closer to flat.}
  \label{fig:aggregate_scaling_2km}
\end{figure}

We have repeated this calculation for the full 36 antenna, 6km ASKAP (see figure \ref{fig:aggregate_scaling_6km}).

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/aggregate_scaling_single_ASKAP_beam_ASKAP_6km_6144.png}
  \caption{Aggregate predicted scaling curve for 6km configuration. Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10, 30], and the overlap is 48 pixels. The image size is 6144 by 6144 pixels.}
  \label{fig:aggregate_scaling_6km}
\end{figure}

If the deconvolution were to be done on the mosaiced data then the full mosaiced primary beam (see figure \ref{fig:full_askap_pb}) would be appropriate. This primary beam is larger and flatter so we would expect that the behaviour would be similar to the no-primary beam approach. The results for mosaiced primary beam are shown in figure \ref{fig:full_askap_aggregate_scaling_2km}. (Compare this to figure \ref{fig:aggregate_scaling_2km}).

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/full_askap_pb_4096.png}
  \caption{Full ASKAP primary formed by summing PAF beams in quadrature. The image size is 4096 by 4096 pixels.}
  \label{fig:full_askap_pb}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/full_askap_aggregate_scaling_ASKAP_2km_4096.png}
  \caption{Aggregate scaling curve for 2km ASKAP configuration using the full mosaic primary beam. Fraction of flux versus rank for various numbers of facets, for realistic image (constructed from S3-SEX) with a primary beam applied and then clipped at 1mJy/beam. Scales are [0, 3, 10], and the overlap is 16 pixels. The image size is 4096 by 4096 pixels.}
  \label{fig:full_askap_aggregate_scaling_2km}
\end{figure}

Figure \ref{fig:number_active_ASKAP_2km} shows the scaling of number of active sub-images with increasing cycle number in major/minor cycles. The processing scaling improves with cycle number.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/number_active_Single_ASKAP_beam_ASKAP_2km_2048.png}
  \caption{Number of active ranks vs major/minor cycle, for 2km ASKAP configuration using the single primary beam. The threshold for each cycle starts at $\sqrt{0.1}$ Jy/beam and proceeds in geometric progression to 0.1 mJy/beam.}
  \label{fig:number_active_ASKAP_2km}
\end{figure}

\subsection{Confirmation of scaling}

Although the above behaviour does need to be verified on a large scale cluster, there is much that can be learned using small scale simulations on an 8 core Linux desktop.

In figure \ref{fig:culm_time_vs_cycle} we show time taken per cycle for a simulation of the 2km ASKAP configuration using the single primary beam. In figure \ref{fig:time_vs_rank} we show the processing time as a function of sub-image for each cycle.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/culm_time_per_cycle.png}
  \caption{Cumulative Measured time taken for the clean of each sub-image vs sub-image rank for each major/minor cycle, for 2km ASKAP configuration using the single primary beam. The threshold for each cycle starts at $\sqrt{0.1}$ Jy/beam and proceeds in geometric progression to 0.1 mJy/beam. The work increases steadily with cycle. Note that there is a stopping problem that can consume extra time, as well as the costs of a major cycle.}
  \label{fig:culm_time_vs_cycle}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{./pngs/time_versus_rank.png}
  \caption{Measured time taken for the clean of each sub-image vs sub-image rank for each major/minor cycle, for 2km ASKAP configuration using the single primary beam. The threshold for each cycle starts at $\sqrt{0.1}$ Jy/beam and proceeds in geometric progression to 0.1 mJy/beam. The work increases steadily with cycle. Note that there is a stopping problem that can consume extra time, as well as the costs of a major cycle.}
  \label{fig:time_vs_rank}
\end{figure}

\pagebreak
\section{Porting to ASKAPsoft}

The sub-image \MAM\ algorithm is now present in the ARL and can be run on both simulated and real data. Porting the algorithm to ASKAPsoft requires implementation in C++. ASKAPsoft has much of the necessary library capabilities, with the exception of sub-image scatter/gather.

The core ASKAPsoft \MAM\ algorithm, DeconvolverMultiTermBasisFunction.tcc, can be retained as is.

@@

For the deconvolution step, it may be convenient to use a separate MPI work group. The ASKAPsoft master/worker framework nominally supports this.

Since the scaling curves are relatively flat (see figures \ref{fig:aggregate_scaling} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing. Load balancing would then only be needed for the high rank tail and should be performed automatically by the master/worker pattern.

The major python-based ARL capabilities that would need to be reproduced in ASKAPsoft are for scatter and gather of an image into sub-images, allowing overlap and tapering.

At the high numbers of processors, reduction in the scatter/gather steps may be needed. The total amount of network traffic is the full image size times the number of Taylor terms. Before beginning the sub-image cleans, the sub-images have to be scattered to the workers.

To facilitate development, the sub-image \MAM\ should first be implemented in cdeconvolver to resolve the issues around use of the master-worker scheme and then ported to cimager.

@@

The work plan to bring the Sub-image \MAM\ to ASKAPsoft. is straightforward. The steps required are:
